
# This file is generating the data files for ML Task 1 and Task 2
# It needs to be called in the folder where ll the 1FBI_XAnalyses/ folders are
# It browses along all the XXXX_XAnalyses/ folders and creates:
#     - datasets for Task 1 for each antigen (inside their respective Analyses folder)
#     - merged datasets along all antigens for ML Task 2
```{r}

library(ggplot2)
library(dplyr)

degree = 1

if(degree == 1){
    folder = "D:/pprobert/Datasets/Features/"
}
if(degree == 2){
    folder = "D:/pprobert/FeaturesDegree2/"
}
```

```{r}
```


```{r}
setwd(folder)

#List of antigens we did process
toParse <- list.dirs(path = ".", full.names = TRUE, recursive = TRUE)
listAvailable <- c()
for (rep in toParse){
    #print(substring(rep, nchar(rep) - 7, nchar(rep)))
    if((nchar(rep) > 1) & (substring(rep, nchar(rep) - 7, nchar(rep)) == "Analyses")){
        #print(rep)
        #Remove ./ and Analyses to get back antigenID
        listAvailable <- c(listAvailable, substring(rep, 3, nchar(rep) - 8))
    }
}
listAvailable    
```

```{r}
length(listAvailable)
```



```{r}

```

```{r}
a = c(8,1,2,3,4,4,5)
duplicated(a)
```

```{r}
#Adds a star to the paratope
Parastarise <- function(x){
    return(paste(x ,"*", sep=""));
}
Parastarise("A1B3C2")
```



```{r}
Parastarization <- function(df){
    df2 <- data.frame(df[,c(1,2,3,4,6,8,10)], apply(df[ , c(5,7,9,11)], 2, Parastarise), df[,c(12,13)])
    return(df2)
}
```



```{r}
#Removes the degree...
library(stringr)
Undeg <- function(x){
    return(str_replace_all(str_replace_all(x, "[12345]", ''), "--","-"));
}
Undeg("A1B3--C2--D3----A1")


```

```{r}
#Generate paratope-epitope map,

#Now we will have two variants: Either we generate datasets of CDR3s,
# or we generate from slices. The source files will be different:
# - antigenID, "_MascotteFeatures.txt" for the CDR3 datasets, then we output CDR3s and the AA usage of CDR3s

#No balancing needed
generateMLTask4 <- function(folderFeatures, antigenID, useSlices=TRUE, uniqueSlices=TRUE, binderRef="Mascotte", nameOut="Task4"){
    
    #Step 1: read filtered files with either the top 1% CDR3s or (if useSlices=true) all the slices that are above this same threshold
    suffix = ""
    if(degree == 2){
        suffix = "D2"
    }
    if(useSlices){
        fileFeaturesBinders = paste(folderFeatures, antigenID, "Analyses/", antigenID, "_", binderRef, "SlicesFeatures", suffix, ".txt", sep="")
    } else {
        fileFeaturesBinders = paste(folderFeatures, antigenID, "Analyses/", antigenID, "_", binderRef, "Features", suffix, ".txt", sep="")
    }
    print(fileFeaturesBinders)
    
    bind <- read.table(fileFeaturesBinders, header=TRUE)
    head(bind)
    
    print(paste(length(bind$ID_slide_Variant), antigenID))
    
    if(useSlices){
        # A duplicate Slice should have
        if(uniqueSlices){
            bind <- bind[!duplicated(bind$Slide),]   
        }
    }
    print(paste(length(bind$ID_slide_Variant), "Binders"))
    
    # Get the distribution of CDR3 size on both binders and nonbinders
    #distribCDR <- bind %>% group_by(sizeCDR3) %>% mutate(number=n()) %>% slice(1) %>% select(sizeCDR3, number)


    # Now shorten the files with only the columns of interest (Slide or CDR3, with the respective AA composition)
    # We keep track of the CDR3 size to later sample from a wished CDR3 lentgth distribution among non-binders (we have a lot of them)
    if(useSlices){
        AllBinders <- cbind(bind) %>% mutate(Label=antigenID) %>% select(Slide, Label, AGboundPositions, hotspot_ID, seqAGEpitope, seqABParatope, motifAGEpitope, motifABParatope, agregatesAGEpitope, agregatesABParatope, chemicalAGEpitope, chemicalABParatope,  motifsSizeGapsLigand, motifsSizeGapsRec, AAcompoFullSlice, sizeCDR3)
       
    } else {
        AllBinders <- cbind(bind) %>% mutate(Label=antigenID) %>% select(CDR3, Label, AGboundPositions, hotspot_ID, seqAGEpitope, seqABParatope, motifAGEpitope, motifABParatope, agregatesAGEpitope, agregatesABParatope, chemicalAGEpitope, chemicalABParatope,  motifsSizeGapsLigand, motifsSizeGapsRec, AAcompoFullCDR, sizeCDR3)
        
    }
    AllBinders <- Parastarization(AllBinders) #Order is changed within cols 5-11 so this should still work:
    AllBindersNoDegree <- data.frame(AllBinders[,c(1,2,3)], apply(AllBinders[ , c(4, 5,6,7,8,9,10,11)], 2, Undeg), AllBinders[,c(12,13)])

    if(useSlices){
        write.table(AllBinders, paste(antigenID, "_", nameOut, "_Slice_ParaEpi", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
        write.table(AllBindersNoDegree, paste(antigenID, "_", nameOut, "_Slice_ParaEpi_NoDeg", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
    } else {
        write.table(AllBinders, paste(antigenID, "_", nameOut, "_CDR3_ParaEpi", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
        write.table(AllBindersNoDegree, paste(antigenID, "_", nameOut, "_CDR3_ParaEpi_NoDeg", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
    }
    return(AllBinders)
}
```

```{r}
antigenID = "1ADQ_A"
#folder = "D:/pprobert/Datasets/Features/"
test <- generateMLTask4(folder, antigenID, useSlices = TRUE, uniqueSlices=TRUE, binderRef="Mascotte", nameOut="Task4")
test
```

```{r}
Parastarization(test)
```

```{r}
regenerateData = FALSE
if(regenerateData){
    #antigenID = "1ADQ_A"
    #Note: now, memory will explode, hihi
    pulledCDR3 <- NULL
    pulledSlices <- NULL
    for(antigenID in listAvailable){
        resCDR <- generateMLTask4(folder, antigenID, useSlices = FALSE)
        resSlice <- generateMLTask4(folder, antigenID, useSlices = TRUE, uniqueSlices = TRUE)
        
    	#resCDRHarderV1 <- generateMLTask1(antigenID, "Binder", useSlices = FALSE, uniqueSlices=FALSE, binderRef="Mascotte", nonbinderRef="LooserExclusive", nameOut="Task1aMvsL")
    	#resSliceHarderV1 <- generateMLTask1(antigenID, "Binder", useSlices = TRUE, uniqueSlices=TRUE, binderRef="Mascotte", nonbinderRef="LooserExclusive", nameOut="Task1aMvsL")
        #resCDRHarderV2 <- generateMLTask1(antigenID, "Binder", useSlices = FALSE, uniqueSlices = FALSE, binderRef="Heroes", nonbinderRef="MascotteExclusive", nameOut="Task1bHvsM")
        #resSliceHarderV2 <- generateMLTask1(antigenID, "Binder", useSlices = TRUE, uniqueSlices = TRUE, binderRef="Heroes", nonbinderRef="MascotteExclusive", nameOut="Task1bHvsM")
    	
        pulledCDR3 <- rbind(pulledCDR3, resCDR)
        pulledSlices <- rbind(pulledSlices, resSlice)
    }
}
```
```{r}
#a <- read.table("D:/pprobert/FeaturesDegree2/4N9G_CAnalyses/4N9G_C_MascotteFeaturesD2.txt")
```

```{r}
regenerateData = FALSE
if(regenerateData){
    head(pulledSlices)
}
```
```{r}
degree
```


```{r}
library(dplyr)
```

```{r}
regenerateData = FALSE
if(regenerateData){
    if(degree == 1){
        write.table(pulledSlices, paste("Task4_Merged_Slice_ParaEpi.txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
        write.table(pulledCDR3, paste("Task4_Merged_CDR3_ParaEpi.txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
    }
    if(degree == 2){
        write.table(pulledSlices, paste("Task4_Merged_Slice_ParaEpiD2.txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
        write.table(pulledCDR3, paste("Task4_Merged_CDR3_ParaEpiD2.txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
    }
}
```




# ===========0 reopening instead of redoing everyh
```{r}
#degree = 2
print(degree)
folderSaved = "D:/pprobert/Datasets/Task4/"
#reprocess <- pulledSlices # read.table(paste(folderSaved, "Task4_Merged_Slice_ParaEpi.txt", sep=""), header=TRUE)

if(degree == 1){
    reprocess <- read.table(paste(folderSaved, "Task4_Merged_Slice_ParaEpi.txt", sep=""), header=TRUE)
    # paste("Task4_Merged_CDR3_ParaEpi.txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
}
if(degree == 2){
    reprocess <- read.table(paste(folderSaved, "Task4_Merged_Slice_ParaEpiD2.txt", sep=""), header=TRUE)
    #write.table(pulledCDR3, paste("Task4_Merged_CDR3_ParaEpiD2.txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
}
```

```{r}
reprocess
```
```{r}
suffix = ""
if(degree == 2){
    suffix = "D2"
}
suffix
```


```{r}
Task4_I_ABDB_Motifs <- reprocess %>% select(motifsSizeGapsLigand, motifsSizeGapsRec) %>% group_by(motifsSizeGapsLigand, motifsSizeGapsRec) %>% mutate(duplicates = n()) %>% ungroup() #slice(1)
Task4_I_ABDB_Motifs
```

```{r}
write.table(Task4_I_ABDB_Motifs, paste("Task4_I_ABDB_Motifs", suffix, "Ungrouped.txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```


```{r}
Task4_A_EpiSeq_ParaSeq <- reprocess %>% select(seqAGEpitope, seqABParatope) %>% group_by(seqAGEpitope, seqABParatope) %>% mutate(duplicates = n()) %>% slice(1)
Task4_A_EpiSeq_ParaSeq
```

```{r}
write.table(Task4_A_EpiSeq_ParaSeq, paste("Task4_A_EpiSeq_ParaSeq", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```




```{r}
Task4_B_EpiMotif_ParaMotif <- reprocess %>% select(motifAGEpitope, motifABParatope) %>% group_by(motifAGEpitope, motifABParatope) %>% mutate(duplicates = n()) %>% slice(1)
Task4_B_EpiMotif_ParaMotif
```

```{r}
write.table(Task4_B_EpiMotif_ParaMotif, paste("Task4_B_EpiMotif_ParaMotif", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```



```{r}
Task4_C_EpiAgr_ParaAgr <- reprocess %>% select(agregatesAGEpitope, agregatesABParatope) %>% group_by(agregatesAGEpitope, agregatesABParatope) %>% mutate(duplicates = n()) %>% slice(1)
Task4_C_EpiAgr_ParaAgr
```

```{r}
write.table(Task4_C_EpiAgr_ParaAgr, paste("Task4_C_EpiAgr_ParaAgr", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```


```{r}
Task4_D_EpiChem_ParaChem <- reprocess %>% select(chemicalAGEpitope, chemicalABParatope) %>% group_by(chemicalAGEpitope, chemicalABParatope) %>% mutate(duplicates = n()) %>% slice(1)
Task4_D_EpiChem_ParaChem
```

```{r}
write.table(Task4_D_EpiChem_ParaChem, paste("Task4_D_EpiChem_ParaChem", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```




```{r}
#head(pulledSlices)
```




```{r}
#head(pulledSlices)
```
```{r}
#Useless, no?
pulledSlices = reprocess
```

```{r}
pulledSlicesNoDegree <- data.frame(reprocess[,c(1,2,3)], apply(reprocess[ , c(4, 5,6,7,8,9,10,11)], 2, Undeg), reprocess[,c(12,13)])
pulledSlicesNoDegree
```
```{r}
#pulledCDR3NoDegree <- data.frame(pulledCDR3[,c(1,2,3)], apply(pulledCDR3[ , c(4, 5,6,7,8,9,10,11)], 2, Undeg), pulledCDR3[,c(12,13)])
#pulledCDR3NoDegree
```

#write.table(pulledCDR3, "Task2A_mergedBalanced.txt", row.names = FALSE, sep = "\t", quote = FALSE)
#write.table(pulledSlice, "Task2A_SlicesmergedBalanced.txt", row.names = FALSE, sep = "\t", quote = FALSE)

```{r}
write.table(pulledSlicesNoDegree, paste("Task4_Merged_Slice_NoDegree_ParaEpi", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
#write.table(pulledCDR3NoDegree, paste("Task4_Merged_CDR3_NoDegree_ParaEpi.txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```




```{r}
Task4_E_EpiSeq_ParaSeq <- pulledSlicesNoDegree %>% select(seqAGEpitope, seqABParatope) %>% group_by(seqAGEpitope, seqABParatope) %>% mutate(duplicates = n()) %>% slice(1)
Task4_E_EpiSeq_ParaSeq
```

```{r}
write.table(Task4_E_EpiSeq_ParaSeq, paste("Task4_E_EpiSeq_ParaSeq_NoDeg", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```




```{r}
Task4_F_EpiMotif_ParaMotif <- pulledSlicesNoDegree %>% select(motifAGEpitope, motifABParatope) %>% group_by(motifAGEpitope, motifABParatope) %>% mutate(duplicates = n()) %>% slice(1)
Task4_F_EpiMotif_ParaMotif
```

```{r}
write.table(Task4_F_EpiMotif_ParaMotif, paste("Task4_F_EpiMotif_ParaMotif_NoDeg", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```



```{r}
Task4_G_EpiAgr_ParaAgr <- pulledSlicesNoDegree %>% select(agregatesAGEpitope, agregatesABParatope) %>% group_by(agregatesAGEpitope, agregatesABParatope) %>% mutate(duplicates = n()) %>% slice(1)
Task4_G_EpiAgr_ParaAgr
```

```{r}
write.table(Task4_G_EpiAgr_ParaAgr, paste("Task4_G_EpiAgr_ParaAgr_NoDeg", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```


```{r}
Task4_H_EpiChem_ParaChem <- pulledSlicesNoDegree %>% select(chemicalAGEpitope, chemicalABParatope) %>% group_by(chemicalAGEpitope, chemicalABParatope) %>% mutate(duplicates = n()) %>% slice(1)
Task4_H_EpiChem_ParaChem
```

```{r}
write.table(Task4_H_EpiChem_ParaChem, paste("Task4_H_EpiChem_ParaChem_NoDeg", suffix, ".txt", sep=""), row.names = FALSE, sep = "\t", quote = FALSE)
```

