```{r}
folder = "D:/pprobert/Datasets/Task1AllAGs/"
folderAnalyses = "D:/pprobert/Datasets/AnalyzedV3/"

#This is the final nonredundant selection of antigens, n=142
lib <- c("1ADQ_A", "1FBI_X", "1FNS_A", "1FSK_A", "1H0D_C", "1JPS_T", "1KB5_AB", "1NCB_N", "1NSN_S", "1OAZ_A", "1OB1_C", "1OSP_O", "1PKQ_J", "1QFW_AB", "1RJL_C", "1S78_B", "1TQB_A", "1WEJ_F", "1YJD_C", "1ZTX_E", "2ARJ_RQ", "2B2X_A", "2FD6_AU", "2HFG_R", "2IH3_C", "2JEL_P", "2Q8A_A", "2R29_A", "2R4R_A", "2R56_A", "2UZI_R", "2VXQ_A", "2VXT_I", "2W9E_A", "2WUC_I", "2XQB_A", "2XWT_C", "2YC1_C", "2YPV_A", "2ZCH_P", "3BGF_S", "3BN9_A", "3CVH_ABC", "3DVG_XY", "3EFD_K", "3GI9_C", "3HI6_A", "3JBQ_B", "3KJ4_A", "3KR3_D", "3KS0_J", "3L5X_A", "3L95_X", "3MJ9_A", "3NCY_A", "3NFP_I", "3NH7_A", "3Q3G_E", "3R08_E", "3R1G_B", "3RAJ_A", "3RKD_A", "3RVV_A", "3SKJ_E", "3SQO_A", "3TT1_A", "3U9P_C", "3UBX_A", "3V6O_A", "3VG9_A", "3VRL_C", "3WD5_A", "4AEI_A", "4CAD_C", "4DKE_A", "4H88_A", "4HC1_B", "4HJ0_B", "4I18_R", "4I77_Z", "4K24_A", "4K3J_A", "4K9E_C", "4KI5_M", "4KXZ_A", "4LU5_B", "4MXV_B", "4N9G_C", "4NP4_A", "4OII_A", "4OKV_E", "4PP1_A", "4QCI_D", "4QEX_A", "4QNP_A", "4QWW_A", "4R9Y_D", "4RGM_S", "4U1G_A", "4U6V_A", "4WV1_F", "4Y5V_C", "4YPG_C", "4YUE_C", "4ZFG_A", "4ZFO_F", "4ZSO_E", "5B8C_C", "5BVP_I", "5C0N_A", "5C7X_A", "5CZV_A", "5D93_A", "5DFV_A", "5DHV_M", "5DMI_A", "5DO2_B", "5E8D_A", "5E8E_LH", "5E94_G", "5EII_G", "5EPM_C", "5EU7_A", "5EZO_A", "5F3B_C", "5FB8_C", "5H35_C", "5HDQ_A", "5HI4_B", "5IKC_M", "5J13_A", "5JW4_A", "5JZ7_A", "5KN5_C", "5KTE_A", "5L0Q_A", "5LQB_A", "5MES_A", "5T5F_A", "5TH9_A", "5TLJ_X", "5TZ2_C");
length(lib)
```

```{r}
#List of antigens we did process, as comparison
setwd(folderAnalyses)
toParse <- list.dirs(path = ".", full.names = TRUE, recursive = TRUE)
libAvailable <- c()
for (rep in toParse){
    #print(substring(rep, nchar(rep) - 7, nchar(rep)))
    if((nchar(rep) > 1) & (substring(rep, nchar(rep) - 7, nchar(rep)) == "Analyses")){
        #print(rep)
        #Remove ./ and Analyses to get back antigenID
        libAvailable <- c(libAvailable, substring(rep, 3, nchar(rep) - 8))
    }
}
libAvailable 
length(libAvailable)
```


```{r}
library(dplyr)
library(ggplot2)
```

#just a test to see formats
```{r}
test <- read.table(paste(folder, lib[1], "_Task1_SlicesBalancedData.txt", sep=""), header=TRUE)
print(length(test[which(test$Label=="Binder"),]$Slide))
print(length(unique(test[which(test$Label=="Binder"),]$Slide)))
print(length(test[which(test$Label=="NonBinder"),]$Slide))
print(length(unique(test[which(test$Label=="NonBinder"),]$Slide)))
print(length(unique(test$Slide)))
```

```{r}
statsData <- NULL
for(antigenID in lib){
    print(antigenID)
    fileBalanced <- paste(folder, antigenID, "_Task1_SlicesBalancedData.txt", sep="")
    test <- read.table(fileBalanced, header=TRUE)
    NB = length(test[which(test$Label=="Binder"),]$Slide)
    NUB = length(unique(test[which(test$Label=="Binder"),]$Slide))
    NNonB = length(test[which(test$Label=="NonBinder"),]$Slide)
    NUNonB = length(unique(test[which(test$Label=="NonBinder"),]$Slide))
    Tot = length(unique(test$Slide))
    newRow <- data.frame(antigenID,  NB, NUB, NNonB, NUNonB, Tot)  
    names(newRow) <- c("Antigen", "NBinders", "NUniqueBinders", "NNonBinders", "NUniqueNonBinders", "Total")  
    statsData <- rbind(statsData, newRow)
}


#https://www.c-sharpcorner.com/article/r-data-frame-operations-adding-rows-removing-rows-and-merging-two-data-frame/
  
```
```{r}
head(statsData)
```

```{r}
statsData %>% arrange(desc(NBinders))
```
```{r}
write.table(statsData, row.names=FALSE, sep="\t")
```


```{r}
library(reshape2)
```


```{r}
d <- melt(statsData, id.vars="Antigen")
d
```
```{r}
library(ggrepel)
```

```{r}
ggplot(d, aes(x=factor(variable),y=value, fill=variable)) + 
  geom_point() + 
  stat_smooth() + theme_classic() + geom_boxplot(position=position_dodge(0.0), width=0.2, outlier.shape = NA)
```

#Here, each file says binder or non binder
#Now I add a column class where it will say "AntigenID" or "Nonbinder". Both columns will be
```{r}
df <- NULL
for(antigenID in lib){
    print(antigenID)
    fileBalanced <- paste(folder, antigenID, "_Task1_SlicesBalancedData.txt", sep="")
    a <- read.table(fileBalanced, header=TRUE)
    b <- a %>% mutate(Class = ifelse(Label == "Binder", antigenID, Label))
    df <- rbind(df, b)
}
df
```

```{r}
print(paste("Full Database task 2 has", length(df$Slide), "sequences"))
print(paste("Full Database task 2 has", length(unique(df$Slide)), "sequences"))

```

```{r}
repeatedSeqs <- df %>% group_by(Slide) %>% mutate(Repeats = n()) %>% slice(1) %>% arrange(desc(Repeats))
head(repeatedSeqs)
```




```{r}
p1 <- ggplot(repeatedSeqs, aes(x=Repeats)) + geom_histogram() + theme_classic() + scale_x_continuous(name="Duplicates", limits=c(0, 30))
p1
```
#If want to separate binders and nonbinders...
```{r}
repeatedSeqsPerLabel <- df %>% group_by(Slide, Label) %>% mutate(Repeats = n()) %>% slice(1) %>% arrange(desc(Repeats))
head(repeatedSeqsPerLabel)

p2 <- ggplot(repeatedSeqsPerLabel, aes(x=Repeats, fill = factor(Label))) + geom_histogram() + theme_classic() + scale_x_continuous(name="Duplicates", limits=c(0, 30))
p2
```
#What should NOT be done. because labels don'tmean anything anymore
```{r}
p2 <- ggplot(repeatedSeqs, aes(x=Repeats, fill = factor(Label))) + geom_histogram() + theme_classic() + scale_x_continuous(name="Duplicates", limits=c(0, 30))
p2
```





```{r}
test <- df[c(1,3,3,10,100,1000, 200002, 200003),]
```


```{r}
test
```
```{r}
newRow <- test[1,]
newRow["Class"] = "1FBI_X"
newRow["Label"] = "Binder"
test[2,"Class"] = "1FBI_X"
test[2,"Label"] = "Binder"
test <- rbind(test, newRow)
test <- rbind(test, test[2,])
test
```

#Now we remove duplicates, only remains antigen => duplicates or nonbinder => duplicates. The nonbinder is not a problem, but duplicates to *different antigens* is a problem.
#It might happen that the same sequences reacts twice to the same antigen, and be careful, we group by label, so loose information. Later would be better to group by Class. (or not do slice)


#What we want: 1/ sequences that bind to two different antigens
```{r}
test %>% select(Slide, Label, Class) %>% arrange(Slide)
```






```{r}
#Think about mutate before slice
groupedDftest <- test %>% group_by(Slide, Label) %>% mutate(Duplicates = n(), nDiffAntigens = n_distinct(Class)) %>% slice(1) %>% arrange(Slide) %>% select(Slide, Label, Duplicates,  nDiffAntigens, Class)
groupedDftest
```




```{r}
groupedDf <- NULL
```




#First, will discard polyreactive: select those that are more than 
```{r}
groupedDfForOutput <- df %>% group_by(Slide, Label) %>% mutate(Duplicates = n(), nDiffAntigens = n_distinct(Class)) %>% arrange(Slide) %>% select(Slide, Label, Duplicates,  nDiffAntigens, Class)
groupedDfForOutput
```
```{r}
write.table(groupedDfForOutput, "Task2Annotated.txt", sep="\t", quote=FALSE, row.names=FALSE)
```


```{r}
groupedDf <- df %>% group_by(Slide, Label) %>% mutate(Duplicates = n(), nDiffAntigens = n_distinct(Class)) %>% slice(1) %>% arrange(Slide) %>% select(Slide, Label, Duplicates,  nDiffAntigens, Class)
groupedDf
```

```{r}
forSaving <- groupedDf %>% filter(Label=="Binder")
```

```{r}
write.table(forSaving, "Task2BindAtLeastOne.txt", quote=FALSE, sep="\t", row.names=FALSE) 
```



```{r}
plotNrBound <- groupedDf %>% filter(Label=="Binder") %>% arrange(desc(Duplicates))
plotNrBound
```




```{r}
makeHisto <- plotNrBound %>% group_by(Duplicates) %>% mutate(Polyreactivity = n()) %>% slice(1)
makeHisto
```
```{r}
makeHisto <- plotNrBound %>% group_by(Duplicates) %>% mutate(Polyreactivity = n()) %>% slice(1)
makeHisto
```


```{r}
ggplot(makeHisto, aes(x=Duplicates, y=Polyreactivity, fill=factor(Label))) + geom_point() + geom_smooth() +  theme_classic() 
```
```{r}
ggplot(makeHisto, aes(x=Duplicates, y=Polyreactivity, fill=factor(Label))) + geom_point() +  theme_classic() + scale_x_continuous(name="Duplicates", limits=c(0, 30))
```


```{r}
ggplot(groupedDf, aes(x=Duplicates, fill=factor(Label))) + geom_histogram() + theme_classic() + scale_x_continuous(name="Duplicates", limits=c(0, 30))
```


```{r}
q <- ggplot(groupedDf, aes(x=Duplicates, fill=factor(Label))) + geom_histogram() + theme_classic() 
```
```{r}
q
```

```{r}
ggplot(makeHisto, aes(x=Duplicates, y=Polyreactivity, fill=factor(Label))) + geom_point() +  theme_classic() + scale_x_continuous(name="Duplicates", limits=c(0, 30))
             
             ggplot(groupedDf, aes(x=Duplicates, fill=factor(Label))) + geom_histogram() + theme_classic() + scale_x_continuous(name="Duplicates", limits=c(0, 30))
```
